--- 
title: ""
pagetitle: "Songs"
site: bookdown::bookdown_site
documentclass: book
link-citations: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, dev.args=list(bg = "transparent"))
```

```{r, results="asis"}
output <- knitr::opts_knit$get("rmarkdown.pandoc.to")# html / latex / docx

if(output == "html"){
  pander::pandoc.header("Songs {-}",1)
  pander::pandoc.image("misc/title.png")
}

```

A collection of hand picked songs. This book is hosted as an online version on https://ratnanil.github.io/Songs/ and includes pdf / epub versions (click on the download symbol). Edits and feedback can be made via the [github repo](https://github.com/ratnanil/songs). The current version was rendered on `r Sys.time()`. 



```{r, include=FALSE}
file.create("docs/.nojekyll")

suppressMessages(library(tabr))
suppressMessages(library(dplyr))
suppressMessages(library(purrr))
suppressMessages(library(ggplot2))
suppressMessages(library(stringr))
suppressMessages(library(gridExtra))
suppressMessages(library(cowplot))
suppressMessages(library(yaml))

test <- FALSE
```


```{r}
# A wrapper around tabr::plot_chord() which allows multiple chords and styles them in my specific way (4 columns, transparent background etc.)
plot_chords <- function(chords){
  # Get the maximum fret-delta of all chords
  delta_max <- map_int(chords,function(x){
    v <- suppressWarnings(as.integer(str_split(x,"",simplify = TRUE)))
    max(v,na.rm = TRUE)-min(v, na.rm = TRUE)
  }) %>%
    max()
  
  # Make a list of all chord plots
  plots <- imap(chords,function(x,y){
    v <- min(suppressWarnings(as.integer(str_split(x,"",simplify = TRUE))),na.rm = TRUE)
    
    plot_chord(x,
               labels = NULL,
               fret_range = c(v,v+delta_max), # so that all chords span the same nr of frets
               point_size = 4,
               label_size = 3) + 
      labs(tag = y) +
      theme(
        panel.background = element_rect(fill = "transparent"), # The theme 
        plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
      )
    })
  chordnames <- map(chords, ~.x[2])
  plot_grid(plotlist = plots,ncol = 4)
}
```


```{r}
# Transform all songs in songs_txt into markdown files


addchord <- function(string,pos,chord){
  paste0(substr(string, 1, pos-1), chord, substr(string, pos, nchar(string)))
}

chordpro_meta <- function(string,metatag){
    regexpression <- paste0("\\{",metatag,":\\s(.+)\\}")
    meta_bool <- str_detect(string,regexpression)
    str_replace(string[meta_bool],regexpression, "\\1")
}


  

  
chordpro_meta_all <- function(string, remove = FALSE){
  chordpro_tags <- c("title", "subtitle", "artist", "composer", "lyricist", "copyright", "album", "year", "key", "time", "tempo", "duration", "capo", "meta")  
  if(!remove){
    out2 <- map(chordpro_tags, ~chordpro_meta(string,.x))
    names(out2) <- chordpro_tags
    out2
  } else{
    remove <- str_detect(string,paste0("(",paste0("\\{",chordpro_tags,":",collapse = "|"),")"))
    string[!remove]
    }
  }

chordpro_environment <- function(string,environment_name){
    start_bool <- str_detect(string,paste0("\\{start_of_",environment_name))
    end_bool <- str_detect(string,paste0("\\{end_of_",environment_name))
    c(which(start_bool),which(end_bool))
}

chordpro_environment_all <- function(string){
    envs <- c("chorus","verse","tab","grid")
    out2 <- map(envs, ~chordpro_environment(string,.x))
    names(out2) <- envs
    out2
    }
```


```{r, eval = FALSE}
dirs <- list.dirs("songs_cho/",full.names = TRUE)
# dir <- dirs[4]
rmd_file_nr = 0

for (file in list.files("songs_md2",pattern = ".Rmd",full.names = TRUE)){file.remove(file)}
for (file in list.files("songs_md3",pattern = ".Rmd",full.names = TRUE)){file.remove(file)}


for (dir in dirs){
  yml <- list.files(dir,pattern = "(.yaml|.yml)",full.names = TRUE)

  
  
  if(length(yml) > 0){
    meta <- read_yaml(yml)
    rmd_file_nr <- rmd_file_nr+1
    rmd_file_nr_chr <- paste0(str_pad(rmd_file_nr,3,pad = "0"),"_")
    
    dir_bn <- basename(dir)
    
    usecss = FALSE
    
    if(usecss){
      fileConn<-file(paste0("songs_md2/",rmd_file_nr_chr,dir_bn,".Rmd"))

    }else{
      fileConn<-file(paste0("songs_md3/",rmd_file_nr_chr,dir_bn,".Rmd"))
    }
    
    paste("#",meta$title) %>%
      writeLines(fileConn)
    
    close.connection(fileConn)
    
    # Todo: include mdchecksums to track changes in files
    
    songs <- list.files(dir, pattern = ".txt",full.names = TRUE)
      
    # gives each file an arbitary number 
    songs_numeric <- map_dbl(basename(songs),~mean(utf8ToInt(.x)))
    names(songs_numeric) <- songs
    songs_numeric <- sort(songs_numeric)
    songs <- names(songs_numeric)
    
    for (song in songs){
  
      song_bn <- basename(song)
      
      song_rl <- song %>%
        readLines(warn = FALSE)
      
      
      if(usecss){
        chordlines_bool <- song_rl %>%
          str_detect("\\[") 
        
        # Take only lines where subsequent line is not a chord line
        chordlines_bool <- chordlines_bool & lead(!chordlines_bool) & lead(str_length(song_rl)>0)
        
        songlines_subsequent_bool <- lag(chordlines_bool,1,default = FALSE)
        
        chordlines <- song_rl[chordlines_bool]
        songlines_subsequent <- song_rl[songlines_subsequent_bool]
      
        songlines_new <- map2_chr(chordlines,songlines_subsequent,function(chordline,songline){
          locs <- str_locate_all(chordline,"\\[\\w+\\]")[[1]]
          # print(locs)
      
          delta_acc = 0
          for(idx in 1:nrow(locs)){
            # print(idx)
            start <- locs[idx,1]
            end <- locs[idx,2]
            chord <- str_sub(chordline,start,end)
            
            start_insert <- start + delta_acc
            if(idx>1){
            } else{
              songline_out <- songline
            }
            songline_out <- addchord(songline_out,start_insert,chord)
            
            delta_acc <- delta_acc+ (end-start+1)
            
          }
          songline_out
        })
        
        song_rl[songlines_subsequent_bool] <- songlines_new
        
        
        song_modified <- song_rl[!chordlines_bool]
        
        
        chordpro_meta_l <- chordpro_meta_all(song_modified)
      
        song_modified <- chordpro_meta_all(song_modified,remove = TRUE)
        
        chordpro_environment_l <- chordpro_environment_all(song_modified) 
        songtitle <- ifelse(length(chordpro_meta_l$artist)>0,paste(chordpro_meta_l$title,chordpro_meta_l$artist, sep = " - "),chordpro_meta_l$title)
          
      
        song_modified[chordpro_environment_l$chorus[1]] <- "**Chorus:**"
        song_modified[chordpro_environment_l$chorus[2]] <- ""
        
        song_modified[chordpro_environment_l$verse[1]] <- ""
        song_modified[chordpro_environment_l$verse[2]] <- ""
        
        song_modified[chordpro_environment_l$tab] <- "```" 
        
        song_modified[chordpro_environment_l$grid] <- "```"
        
        song_modified[chordpro_environment_l$bridge] <- "```"
  
      
        
        # Add the class "show-chords" to lines with chords
        haschords <- str_detect(song_modified,"\\[")
        song_modified[haschords] <- paste0("<p class='song show-chords'>",song_modified[haschords],"</p>")
        
        codechunk <- str_detect(song_modified,"```")
        
        # Add an extra backslash to lines with no chords to force linebreak
        song_modified[!haschords & !codechunk] <- paste0(song_modified[!haschords & !codechunk],"\\")
        
        rmd_file_nr <- rmd_file_nr+1
        rmd_file_nr_chr <- paste0(str_pad(rmd_file_nr,3,pad = "0"),"_")
        
        fileConn<-file(paste0("songs_md2/",paste0(rmd_file_nr_chr, str_replace(song_bn,".txt",".Rmd"))))
        
        song_modified %>%
          prepend(paste("##", songtitle)) %>% # TODO: add other metadata
          str_replace_all("(\\[(\\w+)\\])","<span class='chord'>\\2</span>") %>%
          writeLines(fileConn)
      
        close.connection(fileConn)
      } else{ 
        # do this instead of the fancy css / html stuff
        song_meta <- str_match(song_rl, "\\{(\\w+):\\s(.+)\\}")
        artist_row <- which(song_meta[,2] == "artist")
        artist <- song_meta[artist_row,3]
        title_row <- which(song_meta[,2] == "title")
        title <- song_meta[title_row,3]
        song_rl <- song_rl[-c(artist_row,title_row)]
        
        song_header <- paste("##",title,"-",artist)
        song_rl <- c(song_header,"```",song_rl,"```")
        
        rmd_file_nr <- rmd_file_nr+1
        rmd_file_nr_chr <- paste0(str_pad(rmd_file_nr,3,pad = "0"),"_")
        
        fileConn<-file(paste0("songs_md3/",paste0(rmd_file_nr_chr, str_replace(song_bn,".txt",".Rmd"))))
        
        writeLines(song_rl,fileConn)
      }
      
    }
    
  }
}





```

```{r}

```

